---
title: "Readme"
author: "Alex Franks"
date: "3/28/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

`envelopeR` is the R package associated with the paper ``Reducing Subspace Models for Large-Scale Covariance Regression''.  

**Summary:** We develop an envelope model for joint mean and covariance regression in the large $p$, small $n$ setting.  In contrast to existing envelope methods, which improve mean estimates by incorporating estimates of the covariance structure, we focus on identifying covariance heterogeneity by incorporating information about mean-level differences.  We use a Monte Carlo EM algorithm to identify a low-dimensional
subspace which explains differences in both means and covariances as a function of covariates, and then use MCMC to estimate the posterior uncertainty conditional on the inferred low-dimensional subspace. We demonstrate the utility of our model on a motivating application on the metabolomics of aging.  

Preprint: [https://arxiv.org/abs/2010.00503](https://arxiv.org/abs/2010.00503)

## Installation

```{r, message=FALSE}
devtools::install_github("afranks86/envelopeR")

library(tidyverse)
library(covreg)
library(envelopeR)
library(rstiefel)
library(mvtnorm)
```

## A Test Dataset


```{r generate_test}


test_data <- generate_test_data(q=2, cov_rank=2, intercept=TRUE, gamma_sd=1, error_sd=1, seed=4445)
Y <- test_data$Y
X  <- test_data$X
s  <- test_data$s
p  <- test_data$p
V  <- test_data$V
```

## Fit the Envelope model

```{r fit_envelope}
envfit <- fit_envelope(Y, X, distn="covreg", s=s,
                       Vinit="OLS",
                       verbose_covreg=FALSE)
```


Check inferred subspace is close to the true material subspace (1 means they are identical subspaces).

```{r}
tr(envfit$V  %*% t(envfit$V)  %*% V  %*% t(V))/s
```


Re-run full Bayesian inference conditional on the inferred subspace of material variation.  Alternatively, can get `cov_psamp` from the list returned by `fit_envelope` by running `cov_psamp <- cov.psamp(envfit$covariance_list$covreg_res)`.  

```{r, rerun_mcmc}
YV  <- Y %*% envfit$V
res  <-  covreg::covreg.mcmc(YV ~ X - 1, YV ~ X, niter=10000, nthin=10, verb=FALSE)
cov_psamp  <- covreg::cov.psamp(res)
cov_psamp  <- cov_psamp[, , , 1:1000]
```

## Plotting

### Posterior plot

```{r, posterior_plot, dependson="rerun_mcmc"}
type  <- "mag"

## plot posterior distributions for min, max and quartiles of X
ix <- sort(X[, 1], index.return=TRUE)$ix
obs_to_plot <- ix[c(1, 25, 50, 75, 100)]

## values correspond to the quantiles of x
names(obs_to_plot)  <- c(0, 0.25, 0.5, 0.75, 1)

cols <- colorspace::sequential_hcl(5, "viridis")


post  <- create_plots(envfit$V, cov_psamp,
                      n1=obs_to_plot[1], n2=obs_to_plot[length(obs_to_plot)],
                      to_plot = obs_to_plot, col_values=cols,
                      labels=colnames(Y), plot_type="posterior", alpha=0.5)
post
```

The true eigenvalues and angles on this two-dimensional subspaces are:

```{r, echo=FALSE, true_vals}
rotV  <- rotate_basis(envfit$V, cov_psamp, n1=obs_to_plot[1], n2=obs_to_plot[length(obs_to_plot)])$rotV
R <- t(V) %*% rotV
Vrot  <- V %*% R

evals  <- sapply(1:5, function(i) eigen(test_data$cov_list[[obs_to_plot[i]]])$values[1])
angles  <- sapply(1:5, function(i) {
  evecs  <- eigen(t(R) %*% test_data$cov_list[[obs_to_plot[i]]] %*% R)$vectors
  atan(evecs[2, 1] / evecs[1, 1])
})

tibble(quantile=names(obs_to_plot), eval=evals, angle=angles)
```

### Biplot

```{r, biplot, warn=FALSE}
rownames(envfit$V)  <- 1:p
biplot  <- create_plots(envfit$V, cov_psamp,
                        n1=obs_to_plot[1], n2=obs_to_plot[length(obs_to_plot)],
                        to_plot = obs_to_plot, col_values=cols,
                        labels=colnames(Y), plot_type="biplot")
biplot
```




